Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta-analysis

### 2021-05-12 writing /thesis_draft/_book/thesis.tex
final copy submitted to the library

In addition to the importance of environmental variability across geologic timescales, organisms also interact with environmental variability on several shorter timescales. Organisms and the ecosystems they inhabit have experienced variation on many scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns (e.g.~El Niño Southern Oscillation). As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the dynamics of ecosystems (Khelifa, Blanckenhorn, Roy, Rohner, \& Mahdjoub, 2019; Stewart et al., 2013). How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics. As environmental patterns continue to change, from increased mean temperatures to more severe weather events (e.g.~hurricanes), understanding the limits to organisms' ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Bernardo, 2014; Helmuth et al., 2010; Huey et al., 2012).

@@

There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, e.g.~inflationary effects in sink-source systems (Holt, Barfield, \& Gonzalez, 2003; Vasseur \& Fox, 2009), such that variability aids population density when organisms emigrate. However, there is a lack of consensus as to which scales of biological organization, e.g.~individuals, populations, or communities of organisms, will be most severely affected by different shifts in environmental conditions. We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response. To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.

@@

\caption[Metrics to describe and explain environmental variability.]{An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability. Shown is a comparison of how duration of variability compares within the same environmental factor. Terms highlighted in each panel are described below in the main text}\label{fig:unnamed-chunk-3}
\end{figure}
Environmental variability describes the dynamic characteristics of abiotic and biotic environmental conditions, in addition to how abiotic and biotic factors interact to drive organisms' response to environmental conditions (Gudmundson, Eklöf, \& Wennergren, 2015; Parepa, Fischer, \& Bossdorf, 2013). Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure 1.1).

@@

Aggregating values together, whether it be temperature, light availability, precipitation, allows for a better understanding of trends and patterns as opposed to random events. The mean can be used as a zeroing point for any environmental condition that we can use to compare across patterns, studies, and time. It provides a departure point for what we expect the average value to be for all time. In certain cases, the mean is a useful metric for observing trends by minimizing the noise of environmental conditions, e.g.~averaging population growth rates to understand if a disease is affecting organisms across geographic locations or just specific locations. However, in other cases, the mean is less useful compared to the variability. For instance, if there are key factors contributing to population growth rates from specific locations, comparing the amount of variability in additional variables across locations (individual respiration rates, temperature, etc.) is more fruitful for the task at this scale. At some point however, all data of interest are averages. The important distinction is the scale at which we average the data and the level of detail provided at each scale. Means are more useful in describing variation amongst normally distributed data, however, other statistical metrics like the median or mode may be more appropriate when the data is not normally distributed.

@@

The amount of time an organism spends at different conditions is also important for how we understand and describe environmental variability. Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure 1.1). Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet, Rinehart, Yocum, \& Greenlee, 2018). Performance is a set of whole organism traits (e.g.~assimilation, development, metabolic rate) with the availability to perform dynamics actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Amarasekare \& Savage, 2012; Husak \& Lailvaux, 2017). Studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience warming exposures during a single day. However, there are instances in nature where organisms do experience short periods of extreme temperature changes (e.g.~heat waves, cold fronts, etc.) in which the benefits of FTR in mitigating chill injuries may be helpful. Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with individual, population, and community level processes. For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development (Bowden \& Paitz, 2018; Delmas, Prevot-Julliard, Pieau, \& Girondot, 2007). Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth depending on the magnitude of variance (Lawson, Vindenes, Bailey, \& Pol, 2015). When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (Bastille-Rousseau et al., 2018; Dobramysl \& Tauber, 2013; Romanuk \& Kolasa, 2002).

@@

While not explicitly an area of study for this work, it is important to include how temporal and spatial variability can interact. Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (Farhang-Sardroodi, Darooneh, Kohandel, \& Komarova, 2019; Horne \& Schneider, 1995). Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, e.g.~Diel vertical migration (DVM). In DVM, organisms migrate to the surface and bottom of lakes and oceans based mainly on cues of light availability (Brierley, 2014; Lampert, 1989). However, there are also instances in which the environment is not temporally variable yet an organism utilizes spatial variability to regulate its behavior, e.g.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey, 1974; Kearney, Shine, \& Porter, 2009).

@@

Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (Huey \& Bennett, 1990). Brown et al highlighted the importance of both temperature and body size in the metabolic theory of ecology (MTE) (Brown, Gillooly, Allen, Savage, \& West, 2004). Though highly debated (Clarke, 2006), MTE serves as an origin point for understanding how organisms respond to environmental conditions. MTE is described by the following equation:

@@

Where \(B\) is organismal metabolic rate, \(b_0\) is a normalization constant unrelated to temperate or body size, \(M\) is organismal mass, and \(e^-E/kT\) is referred to as the ``Boltzman factor'', where \(E\) is activation energy, \(k\) is the Boltzman constant and \(T\) is temperature. Together, all of these values aim to describe the relationship between temperature and body size to determine metabolic rate. We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates for larger organisms should be higher and converse for smaller organisms. However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clarke, 2006; Clarke \& Johnston, 1999). Others highlight that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups (Gillooly et al., 2006). Regardless of the debates about the details of MTE, at the very least, it provides a great foundation for thinking about temperature dependent processes.

MTE relies upon temperature impacting metabolic rate, as chemical reactions proceed faster at higher temperatures, which can have downstream effects on fitness characteristics. This is typically described by a characteristic unimodal, right skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
\begin{figure}

@@

\caption[Thermal performance curve components]{Schematic of thermal performance curve and its important components for understanding thermal performance. CTmin denotes Critical Thermal Minimum, CTmax denotes Critical Thermal Maximum, and Topt denotes Thermal Optimum. Adapted from (Krenek et al 2012, Tuff et al 2016).}\label{fig:unnamed-chunk-4}
\end{figure}
The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range. CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, an organism cannot survive (Brown et al., 2004; Dowd, King, \& Denny, 2015). Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.

TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (e.g.~metabolism) (Amarasekare \& Savage, 2012). If only temperature's effect on metabolism were considered, the shape of the relationship may be an exponential one up to a certain point, after whiich it would no longer increase and the organism would fail to survive. However, there are several other constraining factors (e.g.~energy acquisition, nutrient availability, etc.) that drive the characteristically nonlinear pattern of TPCs. TPCs represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.

@@

TPCs have been crucial in predicting, understanding, and capitulating how organisms respond to their environment (Bernhardt, Sunday, Thompson, \& O'Connor, 2018). However, TPCs usually consist of measurements taken across an array of constant temperatures (e.g.~15-35°C in 5°C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance). These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their environments.

Given the characteristically nonlinear shape of TPCs, Jensen's inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions. Jensen's inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel \& Ayres, 1999). For instance, Ruel and Ayres provide the development rate of ectotherms as an example to demonstrate Jensen's inequality (Ruel \& Ayres, 1999). The development rate of ectotherms have a characteristically non-linear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures (Ruel \& Ayres, 1999). With Jensen's inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures (Ruel \& Ayres, 1999). Given this, it is important to recognize that the constant temperatures used to generate TPCs are simply averages of environmental conditions during a moment in time. It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al., 2018; Khelifa et al., 2019).

Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms. For instance, Amarasekare et al (2012) demonstrated the importance where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare \& Savage, 2012; Paaijmans et al., 2013). Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (Amarasekare \& Savage, 2012).

@@

Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level:
\begin{itemize}
\item
  At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, flattening the thermal performance curve (Bartheld, Artacho, \& Bacigalupe, 2017) to changes in TSD ratios (Bowden \& Paitz, 2018)
\item
  At the population level, this usually comes in the form of decreasing population densities (Anders \& Post, 2006)
\item
  At the community level, this usually comes in the form of promoting coexistence in the face of emigration (Descamps-Julien \& Gonzalez, 2005), or population success via inflationary effects (Holt et al., 2003)
\end{itemize}
It is, however, less understood how variation patterns influence these characteristic responses at each level of organization. Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas studies at higher levels of biological organization have explored how colored noise elicits these characteristic responses (Petchey, 2000).

@@

\caption[Meta-analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 19 studies included in the analysis. Red indicates no usable data, magenta indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.}\label{fig:unnamed-chunk-7}
\end{figure}
We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 2.1). After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*)) AND SU=(Life Sciences \& Biomedicine). This search returned a total of 176 results. We downloaded the .bib file of this search and imported this file into the revtools R package (version 0.4.1). Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 2.1). Once we screened studies for relevance to the field of biology and environmental variation, we began excluding studies using the following strict criteria:
\begin{itemize}
\tightlist
\item
  We excluded studies that did not feature both a constant and fluctuating treatment (e.g.~studies with several fluctuating treatments but no constant comparison or studies with a gradient of constant treatment temperatures but no periodic fluctuating treatment)
\item
  We excluded studies that were not completed in a laboratory (e.g.~modeling predictions or results from models, mesocosm experiments, \emph{in situ} experiments)
\item
  We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (e.g.~organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a thermal performance curve (TPC))
\item
  We excluded any articles not written in English
\item
  We excluded studies that did not report measurements of error
\item
  We excluded studies in which the response variables were not comparable or extractable (e.g.~one study performed Principal component analysis for their responses)
\end{itemize}
We excluded studies that did not have the same mean for constant and fluctuating treatments (e.g.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1°C of the two treatments. We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude to most simply variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al., 2004). The final number of data points included in the analysis is broken down by study (Table 2.1).

@@

\caption{\label{tab:unnamed-chunk-9}Covariates included in models}
\centering
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{8cm}}
\toprule
\textbf{Variable} & \textbf{Description}\\
\midrule
\cellcolor{gray!6}{mean\_temp\_constant} & \cellcolor{gray!6}{Reported mean temperature (°C) in study}\\
flux\_range & Reported range (difference between minimum 
                                     and maximum) of temperature fluctuation (°C) experienced\\
org\_level & Level of biological organization, categorized
                                     as either 0 for individual, 1 for population level
\cellcolor{gray!6}{                                     responses, 2 for community level responses}\\
size & Size estimate of study organism, 0 for extra-small (less than 1mm),
                                     1 for small (less than 5 cm), 2 for medium (less than 0.3 m), and 3
                                     for large (greater than 0.3 m)\\
exp\_age & Life stage at which experiment was conducted, 
                                     categorized as 0 for larval, 1 for juvenile (not sexually mature),
\cellcolor{gray!6}{                                     and 2 for adult}\\
\addlinespace
resp\_type & Response variable units classified as rates or traits, 
                                     such that rates included any response variable units with a
                                     “per day” clarifier (e.g. growth rate), all else as traits (e.g. body
                                     width)\\
stressful & Whether or not the organisms experienced thermally stressful 
                                     temperatures throughout study duration, if temperature range 
                                     exceeded Topt, classified as thermally stressful, else not
                                     thermally stressful. See Appendix Table 1 for references in which
\cellcolor{gray!6}{                                     organisms’ thermal stress metrics were included}\\
\bottomrule
\end{tabular}
\end{table}
\clearpage

\hypertarget{metafor}{%
\section{Metafor}\label{metafor}}

@@

The interaction between fluctuation range and mean temperature as well as fluctuation range is best displayed in Figure 3.3. At lower mean temperatures, higher fluctuation ranges generally have a more positive effect on organism responses and performance, though that trend starkly ends at about \(24^{\circ}\)C. However, at higher mean temperatures, lower fluctuation ranges appears to have an equally positive and negative effect on organism responses and performance (Figure 3.3).
\begin{figure}

@@

\caption[Scatterplot of relationship between range, mean, and SMD]{Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and SMD. Temperature fluctuation ranged from 0-20°C, Mean temperature ranged from 7-33°C. SMD ranged from -4 to 6, restricted based on the distribution of SMD to minimize impact of outliers (see Appendix Figure 2).}\label{fig:unnamed-chunk-21}
\end{figure}
The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 3.4). Interestingly, there were not any significant differences between thermal stress (Figure 3.5), life history metric (Figure 3.6), body size (Figure 3.7), or age (Figure 3.8).

>>>>

Organisms and the ecosystems they inhabit have experienced variation on many scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns (e.g.~El Niño Southern Oscillation). As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the dynamics of ecosystems (Khelifa, Blanckenhorn, Roy, Rohner, \& Mahdjoub, 2019; Stewart et al., 2013). How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics. As environmental patterns continue to change, from increased mean temperatures to more severe weather events (e.g.~hurricanes), understanding the limits to organisms' ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Bernardo, 2014; Helmuth et al., 2010; Huey et al., 2012).

@@

There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, e.g.~inflationary effects in sink-source systems (Holt, Barfield, \& Gonzalez, 2003; Vasseur \& Fox, 2009), such that variability aids population density when organisms emigrate. However, there is a lack of consensus as to which scales of biological organization, i.e.~individuals, populations, or communities of organisms, will be most severely affected by different shifts in environmental conditions. We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response. To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.

@@

\caption[Metrics to describe and explain environmental variability.]{An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability. Shown is a comparison of how duration of variability compares within the same environmental factor. Terms highlighted in each panel are described below in the main text.}\label{fig:unnamed-chunk-3}
\end{figure}
Environmental variability describes the dynamic characteristics of abiotic and biotic environmental conditions, in addition to how abiotic and biotic factors interact to drive organisms' response to environmental conditions (Gudmundson, Eklöf, \& Wennergren, 2015; Parepa, Fischer, \& Bossdorf, 2013). Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure 1.1).

@@

Aggregating values together, whether it be temperature, light availability, or precipitation, allows for a better understanding of trends and patterns as opposed to random events. The mean can be used as a zeroing point for any environmental condition that we can use to compare across patterns, studies, and time. It provides a departure point for what we expect the average value to be for all time. In certain cases, the mean is a useful metric for observing trends by minimizing the noise of environmental conditions, e.g.~averaging population growth rates to understand if a disease is affecting organisms across geographic locations or just specific locations. However, in other cases, the mean is less useful compared to the variability. For instance, if there are key factors contributing to population growth rates from specific locations, comparing the amount of variability in additional variables across locations (individual respiration rates, temperature, etc.) is more fruitful for the task at this scale. At some point however, all data of interest are averages. The important distinction is the scale at which we average the data and the level of detail provided at each scale. Means are more useful in describing variation amongst normally distributed data, however, other statistical metrics like the median or mode may be more appropriate when the data is not normally distributed.

@@

The amount of time an organism spends at different conditions is also important for how we understand and describe environmental variability. Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure 1.1). Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet, Rinehart, Yocum, \& Greenlee, 2018). Performance is a set of whole organism traits (e.g.~assimilation, development, metabolic rate) with the availability to perform dynamic actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Amarasekare \& Savage, 2012; Husak \& Lailvaux, 2017). Studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience multiple warming exposures during a single day. However, there are instances in nature where organisms do experience short periods of extreme temperature changes (e.g.~heat waves, cold fronts, etc.) in which the benefits of FTR in mitigating chill injuries may be helpful. Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with individual, population, and community level processes. For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development (Bowden \& Paitz, 2018; Delmas, Prevot-Julliard, Pieau, \& Girondot, 2007). Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth depending on the magnitude of variance (Lawson, Vindenes, Bailey, \& Pol, 2015). When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (Bastille-Rousseau et al., 2018; Dobramysl \& Tauber, 2013; Romanuk \& Kolasa, 2002).

@@

While not explicitly an area of study for this work, it is important to include how temporal and spatial variability can interact. Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (Farhang-Sardroodi, Darooneh, Kohandel, \& Komarova, 2019; Horne \& Schneider, 1995). Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, e.g.~Diel vertical migration (DVM). In DVM, organisms migrate to the surface and bottom of lakes and oceans based mainly on cues of light availability (Brierley, 2014; Lampert, 1989). However, there are also instances in which the environment is not temporally variable, yet an organism utilizes spatial variability to regulate its behavior, e.g.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey, 1974; Kearney, Shine, \& Porter, 2009).

@@

Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (Huey \& Bennett, 1990). Brown et al (2004) highlighted the importance of both temperature and body size in the metabolic theory of ecology (MTE). Though highly debated (Clarke, 2006), MTE serves as an origin point for understanding how organisms respond to environmental conditions. MTE is described by the following equation:

@@

Where \(B\) is organismal metabolic rate, \(b_0\) is a normalization constant unrelated to temperate or body size, \(M\) is organismal mass, and \(e^-E/kT\) is referred to as the ``Boltzman factor'', where \(E\) is activation energy, \(k\) is the Boltzman constant and \(T\) is temperature. Together, all of these values aim to describe the relationship between temperature and body size to determine metabolic rate. We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates for larger organisms should be higher and for smaller organisms should be lower. However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clarke, 2006; Clarke \& Johnston, 1999). Others highlight that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups (Gillooly et al., 2006). Regardless of the debates about the details of MTE, at the very least, it provides a great foundation for thinking about temperature dependent processes.

MTE relies upon temperature impacting metabolic rate, as chemical reactions proceed faster at higher temperatures, which can have downstream effects on fitness characteristics. This is typically described by a characteristic unimodal, left skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
\begin{figure}

@@

\caption[Thermal performance curve components]{Schematic of thermal performance curve and its important components for understanding thermal performance. CTmin denotes Critical Thermal Minimum, CTmax denotes Critical Thermal Maximum, and Topt denotes Thermal Optimum. Adapted from (Krenek et al 2012, Tuff et al 2016).}\label{fig:unnamed-chunk-4}
\end{figure}
The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range. CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, an organism cannot survive (Brown, Gillooly, Allen, Savage, \& West, 2004; Dowd, King, \& Denny, 2015). Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.

TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (e.g.~metabolism) (Amarasekare \& Savage, 2012). If only temperature's effect on metabolism were considered, the shape of the relationship may be an exponential one up to a certain point, after which it would no longer increase and the organism would fail to survive. However, there are several other constraining factors (e.g.~energy acquisition, nutrient availability, etc.) that drive the characteristically nonlinear pattern of TPCs. TPCs represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.

@@

TPCs have been crucial in predicting and understanding how organisms respond to their environment (Bernhardt, Sunday, Thompson, \& O'Connor, 2018). However, TPCs usually consist of measurements taken across an array of constant temperatures (e.g.~15-35°C in 5°C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance). These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their natural environments.

Given the characteristically nonlinear shape of TPCs, Jensen's inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions. Jensen's inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel \& Ayres, 1999). For instance, Ruel and Ayres provide the development rate of ectotherms as an example to demonstrate Jensen's inequality (1999). The development rate of ectotherms has a characteristically nonlinear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures (Ruel \& Ayres, 1999). With Jensen's inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures (Ruel \& Ayres, 1999). Given this, it is important to recognize that the constant temperatures used to generate TPCs are simply averages of environmental conditions during a moment in time. It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al., 2018; Khelifa et al., 2019).

Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms. For instance, Amarasekare et al (2012) demonstrated the importance of where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare \& Savage, 2012; Paaijmans et al., 2013). Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (Amarasekare \& Savage, 2012).

@@

Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level:
\begin{itemize}
\item
  At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, from flattening the thermal performance curve (Bartheld, Artacho, \& Bacigalupe, 2017) to changes in TSD ratios (Bowden \& Paitz, 2018)
\item
  At the population level, this usually comes in the form of decreasing population densities (Anders \& Post, 2006)
\item
  At the community level, this usually comes in the form of promoting coexistence in the face of emigration (Descamps-Julien \& Gonzalez, 2005), or population success via inflationary effects (Holt et al., 2003)
\end{itemize}
It is, however, less understood how variation patterns influence these characteristic responses at each level of organization. Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas studies at higher levels of biological organization have explored how colored noise elicits these characteristic responses (Petchey, 2000).

@@

\caption[Meta-analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 15 studies included in the analysis. Red indicates no usable data, magenta indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.}\label{fig:unnamed-chunk-7}
\end{figure}
We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 2.1). After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*)) AND SU=(Life Sciences \& Biomedicine). This search returned a total of 176 results. We downloaded the .bib file of this search and imported this file into the revtools R package (version 0.4.1). Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 2.1). Once we screened studies for relevance to the field of biology and environmental variation, we began excluding studies using the following strict criteria:
\begin{itemize}
\tightlist
\item
  We excluded studies that did not feature both a constant and fluctuating treatment (e.g.~studies with several fluctuating treatments but no constant comparison or studies with a gradient of constant treatment temperatures but no periodic fluctuating treatment)
\item
  We excluded studies that were not completed in a laboratory (e.g.~modeling predictions or results from models, mesocosm experiments, \emph{in situ} experiments)
\item
  We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (e.g.~organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a TPC)
\item
  We excluded any articles not written in English
\item
  We excluded studies that did not report measurements of error
\item
  We excluded studies in which the response variables were not comparable or extractable (e.g.~one study performed Principal Component Analysis for their responses)
\end{itemize}
We excluded studies that did not have the same mean for constant and fluctuating treatments (e.g.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1°C of the two treatments. We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude most simply to variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al., 2004). The final number of data points included in the analysis is broken down by study (Table 2.1).

@@

\caption{\label{tab:unnamed-chunk-9}Covariates included in models}
\centering
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{8cm}}
\toprule
\textbf{Variable} & \textbf{Description}\\
\midrule
\cellcolor{gray!6}{mean\_temp\_constant} & \cellcolor{gray!6}{Reported mean temperature (°C) in study}\\
flux\_range & Reported range (difference between minimum 
                                     and maximum) of temperature fluctuation (°C) experienced\\
org\_level & Level of biological organization, categorized
                                     as either 0 for individual, 1 for population level
\cellcolor{gray!6}{                                     responses}\\
size & Size estimate of study organism, 0 for extra-small (less than 1mm),
                                     1 for small (less than 5 cm), 2 for medium (less than 0.3 m), and 3
                                     for large (greater than 0.3 m)\\
exp\_age & Life stage at which experiment was conducted, 
                                     categorized as 0 for larval, 1 for juvenile (not sexually mature),
\cellcolor{gray!6}{                                     and 2 for adult}\\
\addlinespace
resp\_type & Response variable units classified as rates or traits, 
                                     such that rates included any response variable units with a
                                     “per day” clarifier (e.g. growth rate), all else as traits (e.g. body
                                     width)\\
stressful & Whether or not the organisms experienced thermally stressful 
                                     temperatures throughout study duration, if temperature range 
                                     exceeded Topt, classified as thermally stressful, else not
                                     thermally stressful. See Appendix Table 1 for references in which
\cellcolor{gray!6}{                                     organisms’ thermal stress metrics were included}\\
\bottomrule
\end{tabular}
\end{table}
\clearpage

\hypertarget{metafor}{%
\section{\texorpdfstring{\emph{Metafor}}{Metafor}}\label{metafor}}

@@

The interaction between fluctuation range and mean temperature as well as effect size is best displayed in Figure 3.3. At lower mean temperatures, higher fluctuation ranges generally have a more positive effect on organism responses and performance, though that trend starkly ends at about \(24^{\circ}\)C. However, at higher mean temperatures, lower fluctuation ranges appear to have an equally positive and negative effect on organism responses and performance (Figure 3.3).
\begin{figure}

@@

\caption[Scatterplot of relationship between range, mean, and SMD]{Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and SMD. Temperature fluctuation ranged from 0-20°C, mean temperature ranged from 7-33°C. SMD ranged from -4 to 6, restricted based on the distribution of SMD to minimize impact of outliers (see Appendix Figure 2).}\label{fig:unnamed-chunk-21}
\end{figure}
The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 3.4). Interestingly, there were not any significant differences between thermal stress (Figure 3.5), life history metric (Figure 3.6), body size (Figure 3.7), or age (Figure 3.8).
